---
title: "Baseball team salaries since 1985"
author: "Chris Hamm"
date: "`r Sys.Date()`"
output: 
  html_document:
  keep_md: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center")
```

### I am learning to use `googleVis` to generate interactive HTML graphics. I thought looking at Major League Baseball team salaries over time would be a fun way to start.

We will use three `packages`:

* `tidyverse` - sort of a supergroup of packages, really
* `Lahman` - which has a `Salary` table
* `googleVis` which will write the HTML code to creat the graphic

Load the libraries:
```{r preliminaries, message = FALSE}
library("tidyverse")
library("Lahman")
library("googleVis")
```

Let's look at the `Salaries` data object:
```{r}
dim(Salaries)
glimpse(Salaries)
```

We can see that the data object contains salaries for every player by year. That isn't really what we want so we need to munge the data a bit.

First, we need to group by `teamID` and `yearID`, then add up all the player salaries for each team by year. I'll also convert the salaries into millions of dollars. 

```{r}
Salary_team_year <- Salaries %>%
  group_by(teamID, yearID) %>%
  summarize(Team_Salary = sum(salary) / 1e6) %>%
  ungroup

glimpse(Salary_team_year)
summary(Salary_team_year)
```

Cool, now our date are set up by team and year. 

Now the `googgleVis` thingy:
```{r}
paid <- gvisMotionChart(Salary_team_year, "teamID", "yearID")
# plot(paid) #uncomment this to run locally
# print(paid, "chart") # this generates the HTML code
```

Have a look at the (admittedly) silly graphic [here](http://pizzabeerandscience.blogspot.com/2017/09/baseball-salaries-by-team-since-1985.html). Change the y-axis to time and color each team uniquely to see how salaries have changed. 
